---
title: Relações entre o uso do espaço, recursos alimentares e concentrações de contaminantes em aves marinhas e estuarinas
subtitle: Projeto de doutorado
author:
  - name: Bruno A Linhares
    email: brunolinhares.bio@gmail.com
    affiliations: 
        - id: some-tech
          name: Universidade Federal do Rio Grande
          department: Laboratório de Aves Aquáticas e Tartarugas Marinhas
          address: Av. Itália s/n
          city: Rio Grande
          state: Rio Grande do Sul
    attributes:
        corresponding: true
  - name: Adalto Bianchini
    email: emaildoadalto@gmail.com
    affiliations:
        - id: another-u
          name: Universidade Federal do Rio Grande
          department: Fisiologia
          address: Um endereço
          city: Rio Grande
          state: Rio Grande do Sul
          postal-code: Zip código
    note: |
      Nota do Adalto
  - name: Leandro Bugoni
    email: leandrobugoni@yahoo.com.br
    affiliations:
        - ref: another-u
    note: Yet another author footnote.
  - name: Derek Zoolander
    email: derek@example.com
    affilations:
        - ref: some-tech
abstract: |
  A poluição é uma das principais formas de impacto humano nos ambientes marinhos. Uma ferramenta para entender a dinâmica da contaminação ambiental é através do estudo de animais indicadores, em especial predadores de topo da cadeia trófica.  Aves marinhas possuem estratégias de forrageio que são variáveis no tempo e no espaço, entre diferentes espécies e indivíduos. Assim, o estudo destas variações, concomitante ao da concentração de contaminantes acumulados nos tecidos das aves, pode permitir compreender os fatores biológicos e ambientais que atuam sobre os níveis de contaminação em predadores de topo da cadeia trófica. O presente projeto, portanto, propõe-se a investigar padrões nas concentrações de elementos-traço e poluentes orgânicos persistentes no sangue e penas de aves marinhas e costeiras amostradas no Banco dos Abrolhos e zona costeira adjacente entre 2006 e 2024, em associação a fatores espaciais e temporais relacionados à história natural, hábito alimentar e uso de habitat. Para isso, dados de contaminação serão utilizados em associação a marcadores de dieta e das áreas de forrageio (isótopos estáveis de carbono, nitrogênio, enxofre e aminoácido-específicos, análise de regurgitados espontâneos e rastreamento remoto). Hipotetiza-se que as concentrações de contaminantes estejam relacionadas a variações temporais, intra e interespecíficas na dieta e ao uso de habitat. Os resultados obtidos com este projeto permitirão avaliar o uso de diferentes espécies de aves marinhas como sentinelas da contaminação ambiental aquática, e os principais fatores e marcadores que influenciam os níveis de contaminantes nas aves.
keywords: 
  - keyword1
  - keyword2
date: last-modified
bibliography: bibliography.bib
format:
  elsevier-pdf:
    keep-tex: true
    journal:
      name: Journal Name
      formatting: preprint
      model: 3p
      cite-style: super
---

Hello there, lets citate something [@Fortunato2010]
Please make sure that your manuscript follows the guidelines in the 
Guide for Authors of the relevant journal. It is not necessary to 
typeset your manuscript in exactly the same way as an article, 
unless you are submitting to a camera-ready copy (CRC) journal.

For detailed instructions regarding the elsevier article class, see   <https://www.elsevier.com/authors/policies-and-guidelines/latex-instructions>

# Bibliography styles

Here are two sample references:  @Feynman @Dirac1953888; [@linhares2023].

By default, natbib will be used with the `authoryear` style, set in `classoption` variable in YAML. 
You can sets extra options with `natbiboptions` variable in YAML header. Example 

```
natbiboptions: longnamesfirst,angle,semicolon
```

There are various more specific bibliography styles available at
<https://support.stmdocs.in/wiki/index.php?title=Model-wise_bibliographic_style_files>. 
To use one of these, add it in the header using, for example, `biblio-style: model1-num-names`.

## Using CSL 

If `cite-method` is set to `citeproc` in `elsevier_article()`, then pandoc is used for citations instead of `natbib`. In this case, the `csl` option is used to format the references. By default, this template will provide an appropriate style, but alternative `csl` files are available from <https://www.zotero.org/styles?q=elsevier>. These can be downloaded
and stored locally, or the url can be used as in the example header.

# Equations

Here is an equation:
$$ 
  f_{X}(x) = \left(\frac{\alpha}{\beta}\right)
  \left(\frac{x}{\beta}\right)^{\alpha-1}
  e^{-\left(\frac{x}{\beta}\right)^{\alpha}}; 
  \alpha,\beta,x > 0 .
$$

Inline equations work as well: $\sum_{i = 2}^\infty\{\alpha_i^\beta\}$

# Figures and tables

@fig-meaningless is generated using an R chunk.

```{r}
#| label: fig-meaningless
#| fig-cap: A meaningless scatterplot
#| fig-width: 5
#| fig-height: 5
#| fig-align: center
#| out-width: 50%
#| echo: false
plot(runif(25), runif(25))
```

# Tables coming from R

Tables can also be generated using R chunks, as shown in @tbl-simple example.

```{r}
#| label: tbl-simple
#| tbl-cap: Caption centered above table
#| echo: true
knitr::kable(head(mtcars)[,1:4])
```

# References {-}
